---
title: "ibutton_Temp"
author: "Isabella Ragonese"
date: "2024-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load libraries
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(rstatix)
library(ggpubr)
library(lubridate)
library(Hmisc)
library(lme4)
library(glmmTMB)
library(sjPlot)
library(car)
library(MuMIn)
library(MASS)
```

## Temp data
      
#read in raw data with ambient and elevated ibutton readings
```{r}
Ambient<-read.csv("AverageAmbient_TEMP2.csv")
Elevated<-read.csv("AverageElevated_TEMP.csv")
```

#transform dataframes to long format
```{r}
A1<-mutate(Ambient, Temp_treatment = "Ambient")
E1<-mutate(Elevated, Temp_treatment = "Elevated")
E1$Stdev = E1$StandardDev

A1long<-pivot_longer(A1,
     cols = c("H1","H6","H7","H8", "H11","H12", "T1","T2","T3","T11","T12","T13","T14","T15","T17" ),
     names_to = "ibuttonID",
     values_to ="Temp")

E1long<-pivot_longer(E1,
                     cols = c("H2","H3","H4","H5","H9","H10","T4","T5","T6","T8","T9","T16","T18","T19","T20"),
                     names_to = "ibuttonID",
                     values_to ="Temp")
A1long<-A1long[,-4]
A1long<-A1long[,-4]
E1long<-E1long[,-4]
E1long<-E1long[,-4]
E1long<-E1long[,-5]

A1_E1_long<-rbind(A1long,E1long)

A1_E1_long$Date<-as.Date.character(A1_E1_long$Date.Time, tryFormats ="%m/%d/%Y %H:%M")

#day vs. night temps 
dates<-A1_E1_long$Date.Time
dates2 <- as.POSIXct(dates, format = "%m/%d/%Y %H:%M")
A1_E1_long$Hour<-format(dates2, format="%H")
A1_E1_long$Hour<-as.numeric(A1_E1_long$Hour)

A1_E1_long$DayNight<-ifelse((A1_E1_long$Hour>6 & A1_E1_long$Hour<20),"Day","Night") #7am-7pm

Am_El_day<-A1_E1_long %>% filter(DayNight=="Day")
Am_El_night<-A1_E1_long %>% filter(DayNight=="Night")

```

#remove outliers from ibutton columns (or temp in Am_El_day and night dfs)
```{r}
outliers<-identify_outliers(as.data.frame(A1_E1_long$Temp))
outliers
#100 outliers; minimum = 41.572

#A1_E1_long_subset<-subset(A1_E1_long, Temp < 41.572)

#no H2
#A1_E1_long_noH2<-A1_E1_long %>% filter(ibuttonID != "H2_temp")

#outliers<-identify_outliers(as.data.frame(A1_E1_long_noH2$Temp))
#outliers
#7 ext outliers; minimum = 42.086

#A1_E1_long_noH2_test<-subset(A1_E1_long_noH2, Temp < 42.086) #eliminates 12 measurements above 40C

#A1_E1_long<-A1_E1_long_noH2_test

A1_E1_long_subset<-subset(A1_E1_long, Temp < 42.086)

A1_E1_long<-A1_E1_long_subset

Am_El_day<-A1_E1_long %>% filter(DayNight=="Day")
Am_El_night<-A1_E1_long %>% filter(DayNight=="Night")

```

#plot difference in mean daily temp
```{r}
#A1_E1_long_noH2<-A1_E1_long %>% filter(ibuttonID != "H2_temp")

Treat_Date_avgtemps<-A1_E1_long %>%
  group_by(Temp_treatment, Date) %>%
  get_summary_stats(Temp, type = "mean_sd")

Treat_Date_avgtemps$se=Treat_Date_avgtemps$sd/sqrt(Treat_Date_avgtemps$n)

Treat_Date_avgtemps=filter(Treat_Date_avgtemps,Date!="2021-06-14") #pre-field date

ggplot(Treat_Date_avgtemps, aes(x=Date,y=mean))+
  geom_line(aes(color=Temp_treatment))+
  geom_ribbon(aes(y=mean, ymin=mean-se, ymax=mean+se, fill=Temp_treatment), alpha=0.2)+
  theme_bw()+
  labs(x="Date",y="mean daily temperature (+/- standared error)")+
  scale_color_manual(values = c("blue","goldenrod"))+
  scale_fill_manual(values = c("blue","goldenrod"))

ggplot(Treat_Date_avgtemps, aes(x=Date,y=mean,color=Temp_treatment))+
  geom_line()+
  geom_ribbon(aes(y=mean, ymin=mean-sd, ymax=mean+sd, fill=Temp_treatment), alpha=0.2)+
  labs(x="Date",y="mean daily temperature (+/- standared dev)")+
  theme_bw()+
  scale_color_manual(values = c("blue","goldenrod"))+
  scale_fill_manual(values = c("blue","goldenrod"))
```

#create df with mean and sd temp for each ibutton (full experiment)
```{r}

per_ibutton_avgtemps_full<-A1_E1_long %>%
  group_by(ibuttonID, Temp_treatment) %>%
  get_summary_stats(Temp, type = "mean_sd")


per_ibutton_avgtemps_full$se=per_ibutton_avgtemps_full$sd/sqrt(per_ibutton_avgtemps_full$n)

#won't work without H2
#per_ibutton_avgtemps_full$PlotNum<-c(8,15,12,14,9,13,24,23,20,6,31,19,16,29,2,18,25,4,32,27,17,30,10,21,22,26,28,3,11,5)

```

#bar plot per ibutton of mean temp
```{r}
button_order<-c("T12","T15","H7","H1","T2","H11","H12","T1","T13","H6","T3","T14","T17","T11","H8","T6","T9","H2","T8","H3","H10","T18","H9","T20","H5","H4","T4","T5","T19","T16")

ggplot(per_ibutton_avgtemps_full)+
geom_bar(aes(x = factor(ibuttonID, button_order), y = mean, 
             fill = Temp_treatment), stat = "identity", show.legend = FALSE)+
  geom_errorbar(aes(x=ibuttonID, y=mean, ymin=mean-se, ymax=mean+se), width=0.5)+
  labs(x="ibutton",y="mean temperature")+
  scale_fill_manual(values = c("blue","goldenrod"))

```

#create df for mean and sd temp for each DAY
```{r}
per_ibutton_Date_avgtemps<-A1_E1_long %>%
  group_by(ibuttonID, Temp_treatment, Date) %>%
  get_summary_stats(Temp, type = "mean_sd")
  

per_ibutton_Date_avgtemps$se=per_ibutton_Date_avgtemps$sd/sqrt(per_ibutton_Date_avgtemps$n)


per_ibutton_Date_avgtemps=filter(per_ibutton_Date_avgtemps,Date!="2021-06-14")

# per_ibutton_Date_avgtemps=filter(per_ibutton_Date_avgtemps,ibutton_ID!="H2")
# per_ibutton_Date_avgtemps=filter(per_ibutton_Date_avgtemps,ibutton_ID!="H1")

ggplot(per_ibutton_Date_avgtemps, aes(x=Date,y=mean,group=ibuttonID,color=Temp_treatment))+
  geom_line()+
  #geom_ribbon(aes(y=mean, ymin=mean-se, ymax=mean+se, fill=Temp_treatment), alpha=.2)+
  labs(x="Date",y="mean temperature")+
  scale_color_manual(values = c("blue","goldenrod"))

```

#repeated measures anova compare ambient and elevated mean daily temp accounting for ibuttonID
http://agroninfotech.blogspot.com/2020/06/two-way-repeated-measures-analysis-in-r.html  
```{r}

#per_ibutton_Date_avgtemps_02<-filter(per_ibutton_Date_avgtemps, ibutton_ID!="H1"&ibutton_ID!="H2") #14 ibuttons in each treatment

#per_ibutton_Date_avgtemps_02

pwc <- per_ibutton_Date_avgtemps %>%
group_by(Date) %>%
pairwise_t_test(
mean ~ Temp_treatment, paired = TRUE,
p.adjust.method = "bonferroni"
)
data.frame(pwc) #each day, mean temp in ambient plots is different from elevated

```

#try anova with A vs. E daytime temps

**excluding ibutton H2 due to consistent higher temperature readings**

```{r}

#Am_El_day = Am_El_day %>% separate(col=ibuttonID,into = c("ibutton_ID", "lag","relative"), "_")
#Am_El_day <- filter(Am_El_day, ibutton_ID!="H1"&ibutton_ID!="H2")
Am_El_day <- filter(Am_El_day, ibuttonID!="H2")

model.aov <- aov(Temp ~ Temp_treatment*Date + Error(ibuttonID/(Date)), data=Am_El_day)
summary(model.aov)

#Am_El_night = Am_El_night %>% separate(col=ibuttonID,into = c("ibutton_ID", "lag","relative"), "_")
#Am_El_night <- filter(Am_El_night, ibutton_ID!="H1"&ibutton_ID!="H2")
Am_El_night <- filter(Am_El_night, ibuttonID!="H2")


model.aov <- aov(Temp ~ Temp_treatment*Date + Error(ibuttonID/(Date)), data=Am_El_night)
summary(model.aov)

meantempday<-Am_El_day %>% group_by(Temp_treatment) %>% 
  summarise(meanTemp=mean(Temp),
            seTemp=sd(Temp)/sqrt(n()))

meantempnight<-Am_El_night %>% group_by(Temp_treatment) %>% 
  summarise(meanTemp=mean(Temp),
            seTemp=sd(Temp)/sqrt(n()))

model.aov <- aov(mean ~ Temp_treatment*Date + Error(ibuttonID/(Date)), data=per_ibutton_Date_avgtemps)
summary(model.aov)

```

