---
title: "MW_warming_Nov_2024"
author: "Isabella Ragonese"
date: "2024-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Description: Data analysis for a 2021 field experiment exploring the effects of elevated temperature, milkweed host plant species, and the interaction between milkweed and temperature on monarch traits and infection.

##inputs: 
# temperature data: "per_ibutton_avgtemps_full.csv" ; 
# monarch data: "MWwarming_comp_May16.csv"
# plant data: "UGA_Cardenolide_Analysis_March2023_tropical.csv"

##outputs: tables and figures in the main text and supplement

##Note: Temperature and humidity data analysis completed in separate rmd files: ibutton_temp.rmd, ibutton_humidity.rmd
# inputs:Temperature data: "AverageAmbient_TEMP2.csv" "AverageElevated_TEMP.csv" ; Hum idity data: "AverageAmbient_RelHum.csv" "AverageElevated_RelHum.csv" ;


#load libraries
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(rstatix)
library(ggpubr)
library(lubridate)
library(Hmisc)
library(lme4)
#library(glmmTMB)
library(sjPlot)
library(car)
library(MuMIn)
#library(MASS)
library(emmeans)
library(DescTools)
#library(performance)
library(lmtest)
library(scales)
```

# functions
```{r, include=FALSE}
#Bolker et al. overdispersion function

overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

```

#########################################################################

## Read in data for monarch, OE and milkweed traits
```{r, include=FALSE, echo=FALSE}
#main dataset
mw2021<-read.csv("MWwarming_comp_May16.csv")

#cardenolide measures
cardenolides<-read.csv("UGA_Cardenolide_Analysis_March2023_tropical.csv")#

cardenolides1<-mw2021 %>% dplyr::select(Temp,Plant_ID) %>% left_join(cardenolides,mw2021,by="Plant_ID")
#monarch data with complete cardenolide data
card2<-distinct(cardenolides1) %>% filter(!is.na(total.card.conc))

#subsets
mw.inf <- subset(mw2021, InfectionStatus==1)
mw.inoc <- subset(mw2021, OE_treatment=="infected")
#comprehensive dataframe with just control individuals
mw.control <- subset(mw2021, OE_treatment=="control", InfectionStatus!=1) 
mw.trop <- subset(mw2021, Milkweed=="tropical")
mw.swamp <- subset(mw2021, Milkweed=="swamp")

#read in daily average temps for each ibutton
per_ibutton_avgtemps_full<-read.csv("per_ibutton_avgtemps_full.csv")

#join avg temp of plot to main dataframe with per_ibutton_avgtemps_full df
mw2021_join<-left_join(mw2021, per_ibutton_avgtemps_full, by="PlotNum")#add overall mean temp

card2$PlotNum<-card2$plot.num
card3<-left_join(card2, per_ibutton_avgtemps_full, by="PlotNum")#add overall mean temp to cardenolide data

##levels of factors
options(contrasts=c("contr.treatment","contr.poly"))
contr.treatment(unique(mw2021$Milkweed))
contr.treatment(unique(mw2021$Temp))


```

################################################
**fitness metrics**

## Development to pupation
```{r}
pupdev_table<-mw2021 %>% 
  group_by(Plant_ID) %>% 
  dplyr::summarise(mean_dev_time = mean(Inoc_to_pupa, na.rm=T),
                   Temp = first(Temp),
                   Milkweed = first(Milkweed),
                   Lineage = first(Lineage),
                   OE_strain = first(OE_strain),
                   OE_treatment = first(OE_treatment))

m0<-lm(log10(mean_dev_time) ~ 1, data = pupdev_table) #null hypothesis
m1<-lm(log10(mean_dev_time) ~ Temp + Milkweed + OE_treatment, data = pupdev_table) #time to pupation depends on main effects of temperature, milkweed sp, and/or inoculation status
m2<- lm(log10(mean_dev_time) ~ Temp*Milkweed + OE_treatment, data = pupdev_table) #interactions between temp and milkweed and/or inoculation and milkweed explain time to pupation

AICc(m0,m1,m2)

summary(m1)
#shorter development times for monarchs on swamp or in ambient plots 

#check normality and variance
plot(m1)
shapiro.test(resid(m1)) #normal residuals
bptest(m1) #homogeneity
overdisp_fun(m1) #not overdispersed

tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

# ggplot(pupdev_table, aes(x=factor(Temp), y=log10(mean_dev_time), fill=Milkweed)) + 
#   geom_boxplot(alpha=0.5, outliers=FALSE)+ #outliers included in jitter
#   geom_jitter(position = position_jitterdodge(jitter.height = 0.01, jitter.width = 0.3), alpha=0.7, aes(colour=Milkweed, shape=Milkweed))+
#   labs(fill = "Milkweed", x = "temperature treatment", y = "time to pupation (log10(days))")+
#   theme_bw()+
#   scale_fill_manual(values = c("darkgreen","darkorange"))+
#   scale_colour_manual(values = c("darkgreen","darkorange"))

#use scale_y_log10 for y axis

ggplot(pupdev_table, aes(x=factor(Temp), y=(mean_dev_time), fill=Milkweed)) + 
  geom_boxplot(alpha=0.5, outliers=FALSE)+ #outliers included in jitter
  geom_jitter(position = position_jitterdodge(jitter.height = 0.01, jitter.width = 0.3), alpha=0.7, aes(colour=Milkweed, shape=Milkweed))+
  labs(fill = "Milkweed", x = "temperature treatment", y = "time to pupation (days)")+
  theme_bw()+
  scale_y_log10(
    breaks = c(10, 12, 14, 16, 18, 20),
    labels = c("10", "12", "14", "16", "18", "20")
  ) +
  scale_fill_manual(values = c("darkgreen","darkorange"))+
  scale_colour_manual(values = c("darkgreen","darkorange"))




```

## Survival probability
```{r}

m0<-glm(Surv_adult ~ 1, data = mw2021, family = binomial)
m1<-glm(Surv_adult ~ Temp + Milkweed + OE_treatment, data = mw2021, family = binomial)
m2<-glm(Surv_adult ~ Temp*Milkweed + OE_treatment, data = mw2021, family = binomial)


AICc(m0,m1,m2)

#model m2 with interaction is slightly better

tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

summary(m2)#temp*milkweed near signif. --> post hoc test for estimated marginal means emmeans?
exp(coefficients(m2))

#calculate R-squared for model
PseudoR2(m2,which="McFadden") #0.073
PseudoR2(m2,which="Nagelkerke") #0.094
PseudoR2(m2,which="Tjur")#0.041 - in table from tab_model()

em <- emmeans(m2, specs ="Milkweed", by ="Temp")
contrast(em, "pairwise", adjust = "Tukey")

require("ggplot2")
emmip(m2, Milkweed ~ Temp, CIs = TRUE, type = "response") +
  theme_bw()+
  scale_colour_manual(values = c("darkgreen","darkorange"))
    #geom_point(aes(x = Temp, y = Surv_adult), data = mw2021, pch = 2, color = "blue")


#########################
#plot
survadult <-  #table of survival to eclosion by treatment 
  mw2021  %>% 
  group_by(Temp, Milkweed, OE_treatment) %>% 
  dplyr::summarise( propsurv_all= sum(Surv_adult, na.rm=T)/n(),
                    seprop_all = sqrt(propsurv_all*(1-propsurv_all)/n()))

ggplot(survadult, aes(x=Temp, y=propsurv_all)) + 
  theme_bw()+
  geom_errorbar(data = survadult, aes(ymin=propsurv_all-seprop_all, 
                                       ymax=propsurv_all+seprop_all,
                                                colour=Milkweed,shape=OE_treatment),width=0.1,
                position=position_dodge(0.4))+
  geom_point(aes(Temp,propsurv_all, colour=Milkweed,shape=OE_treatment),
               position=position_dodge(0.4), cex=3)+
  labs(fill = "Milkweed", x = "temperature treatment", y = "survival probability")+
  scale_fill_manual(values = c("#4E6358","#FC923A"))+
  scale_colour_manual(values = c("#4E6358","#FC923A"))

#not showing OE treatment
survadult <-  #table of survival to eclosion by treatment 
  mw2021  %>% 
  group_by(Temp, Milkweed) %>% 
  dplyr::summarise( propsurv_all= sum(Surv_adult, na.rm=T)/n(),
                    seprop_all = sqrt(propsurv_all*(1-propsurv_all)/n()))

ggplot(survadult, aes(x=Temp, y=propsurv_all)) + 
  theme_bw()+
  geom_errorbar(data = survadult, aes(ymin=propsurv_all-seprop_all, 
                                       ymax=propsurv_all+seprop_all,
                                                colour=Milkweed),width=0.1,
                position=position_dodge(0.4))+
  geom_point(aes(Temp,propsurv_all, colour=Milkweed, shape=Milkweed),
               position=position_dodge(0.4), cex=3)+
  labs(fill = "Milkweed", x = "temperature treatment", y = "survival probability")+
  scale_fill_manual(values = c("darkgreen","darkorange"))+
  scale_colour_manual(values = c("darkgreen","darkorange"))


```

## Pupal mass
```{r}

m0<-lm(PupalMass ~ 1, data = mw2021)
m1<-lm(PupalMass ~ Temp + Milkweed + OE_treatment, data = mw2021)
m2<-lm(PupalMass ~ Temp*Milkweed + OE_treatment, data = mw2021)


AICc(m0,m1,m2)

summary(m2) #infection status non-significant in predicting pupal mass, R2 low
tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

#check normality and variance
plot(m2)
shapiro.test(resid(m2)) #normal residuals
bptest(m2) #homogeneity
overdisp_fun(m2) #not overdispersed

ggplot(mw2021, aes(x=Milkweed, y=PupalMass, fill=OE_treatment)) + 
  geom_boxplot(notch = TRUE, alpha=0.6)+
  geom_jitter(position = position_jitterdodge(jitter.height = 0.01, jitter.width = 0.3), alpha=0.7, aes(colour=OE_treatment))+
  labs(fill = "OE treatment", x = "Milkweed", y = "Pupal mass (g)")+
  scale_fill_manual(values = c("gray","brown"))+
  scale_colour_manual(values = c("gray","brown"))+
  theme_bw()


ggplot(mw2021, aes(x=OE_treatment, y=PupalMass, fill=OE_treatment)) + 
  geom_boxplot(alpha=0.6)+
  geom_jitter(position = position_jitterdodge(jitter.height = 0.01, 
              jitter.width = 0.3), alpha=0.7, aes(colour=OE_treatment))+
  labs(x = "OE", y = "Pupal mass (g)")+
  scale_fill_manual(values = c("gray","brown"))+
  scale_colour_manual(values = c("gray","brown"))+
  theme_bw()

```

## forewing length
```{r}
mw2021_adult<-filter(mw2021, mw2021$Surv_adult==1)

m0<-lm(ForewingLength.mm. ~ 1, data = mw2021_adult)
m1<-lm(ForewingLength.mm. ~ Temp + Milkweed + OE_treatment + Sex, data = mw2021_adult)
m2<-lm(ForewingLength.mm. ~ Temp*Milkweed + OE_treatment + Sex, data = mw2021_adult)


AICc(m0,m1,m2)#m1 with no interaction better than null

summary(m1) #sex, OE treat, milkweed all matter
tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

#check normality and variance
plot(m1)
shapiro.test(resid(m1)) #
bptest(m1) #homogeneity
overdisp_fun(m1) #overdispersed

ggplot(mw2021_adult, aes(x=Sex, y=ForewingLength.mm., fill=Milkweed)) + 
  geom_boxplot()+
  labs(fill = "Milkweed", x = "Infection treatment", y = "Forewing length (mm)")+
  scale_fill_manual(values = c("darkgreen", "orange"))+ theme_bw()

ggplot(mw2021_adult, aes(x=OE_treatment, y=ForewingLength.mm., fill=OE_treatment)) + 
  geom_boxplot()+
  labs(fill = "OE_treatment", x = "Infection treatment", y = "Forewing length (mm)")+
  scale_fill_manual(values = c("gray", "brown"))

```

#Adult longevity (starvation resistance)
```{r}
#we know that infection decreases adult longevity, but do elevated temperatures or either of the milkweed species impact the difference between healthy and infected?

m0<-glm.nb(AdultLongevity ~ 1, data = mw2021_adult) #null
m1<-glm.nb(AdultLongevity ~ OE_treatment + PupalMass + Sex + Temp + Milkweed, data = mw2021_adult) #variables that could impact longevity
m2<-glm.nb(AdultLongevity ~ OE_treatment*Temp + Milkweed + PupalMass + Sex, data = mw2021_adult)
m3<-glm.nb(AdultLongevity ~ OE_treatment*Milkweed + Temp + PupalMass + Sex, data = mw2021_adult)


#model with OE_treatment * temp interaction is best
AICc(m0,m1,m2,m3)

summary(m2) #OE treatment, and the interaction w/ temp significant - elevated temp exacerbates infection cost to longevity
tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

#model check
overdisp_fun(m2) #-->looks ok (poisson was overdispersed)
shapiro.test(resid(m2))

####################################
ggplot(mw2021_adult, aes(x=OE_treatment, y=AdultLongevity, fill=Temp)) + 
  geom_boxplot(alpha=0.5)+
  labs(fill = "Temp", x = "infection treatment", y = "adult starvation resistance (longevity in days))")+
  geom_jitter(position = position_jitterdodge(jitter.height = 0.01, 
              jitter.width = 0.25), alpha=0.7, aes(colour=Temp, shape=Temp))+
  theme_bw()+
  scale_fill_manual(values = c("#26185F", "#18BDB0"))+
  scale_colour_manual(values = c("#26185F","#18BDB0"))+
  scale_shape_manual(values = c(25,22))

ggplot(mw2021_adult, aes(x=OE_strain, y=AdultLongevity, fill=Milkweed)) + 
  geom_boxplot()+
  labs(fill = "Milkweed", x = "infection treatment", y = "adult starvation resistance (longevity in days))")+
  theme_bw()+
  scale_fill_manual(values = c("darkgreen", "orange"))
```

#####################################################
**infection outcomes**

#infection probability
```{r}
#strains used in each treatment
oestrain_treat<-mw.inoc %>% 
  group_by(Temp, Milkweed, OE_strain) %>% 
  summarise(OE_st_ct<-n()) 

mw.inoc1<-filter(mw.inoc, !is.na(PupalMass))
mw.inoc1<-filter(mw.inoc1, !is.na(Sex))
mw.inoc1$Milkweed<-as.factor(mw.inoc1$Milkweed)
mw.inoc1$Temp<-as.factor(mw.inoc1$Temp)

m0<-glm(InfectionStatus ~ 1, data = mw.inoc1, family = binomial)
m1<-glm(InfectionStatus ~ Temp + Milkweed + OE_strain + Sex + PupalMass, data = mw.inoc1, family = binomial)
m2<-glm(InfectionStatus ~ Temp*Milkweed + Sex + PupalMass + OE_strain, data = mw.inoc1, family = binomial)

AICc(m0,m1,m2)#model with main effects better than null

summary(m1)

tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE) #OE strain matters, and temp is marginally non-significant (p= 0.056)

#calculate McFadden's or Tjur's R-squared for model
PseudoR2(m1,which="McFadden") #0.104
PseudoR2(m1,which="Nagelkerke") #0.154
PseudoR2(m1,which="Tjur")#0.111 - in table

#model with interaction
em <- emmeans(m2, specs ="Temp", by ="Milkweed")
contrast(em, "pairwise", adjust = "Tukey")
#model without interaction
em <- emmeans(m1, ~ Temp)
pairs(em, simple = "Temp")

require("ggplot2")
emmip(m2, Milkweed ~ Temp, CIs = TRUE, type = "response") +
  theme_bw()+
  scale_colour_manual(values = c("darkgreen","orange"))


#ggplot
infsuccess <-  #table of infection success by treatment 
  mw.inoc1  %>% 
  group_by(Temp, Milkweed) %>% 
  dplyr::summarise(propinf_all= sum(InfectionStatus, na.rm=T)/n(),
  seprop_all = sqrt(propinf_all*(1-propinf_all)/n()))

infsuccess$Temp<-as.factor(infsuccess$Temp)

#of those inoculated what proportion became infected by treatment?

# ggplot(infsuccess, aes(x=Temp, y=propinf_all)) + 
#   theme_bw()+
#   geom_errorbar(data = infsuccess, aes(ymin=propinf_all-seprop_all, 
#                                        ymax=propinf_all+seprop_all, colour=Milkweed),
#                 width=0.1, position=position_dodge(0.2))+
#   geom_dotplot(aes(fill=Milkweed), binwidth = 0.03, binaxis='y', stackdir='center',dotsize = 0.5,
#                position=position_dodge(0.2))+
#   labs(fill = "Milkweed", x = "temperature treatment", y = "infection probability")+
#   scale_fill_manual(values = c("darkgreen","darkorange"))+
#   scale_colour_manual(values = c("darkgreen","darkorange"))

ggplot(infsuccess, aes(x=Temp, y=propinf_all)) + 
  theme_bw()+
  geom_errorbar(data = infsuccess, aes(ymin=propinf_all-seprop_all, 
                                       ymax=propinf_all+seprop_all, colour=Milkweed),
                width=0.1, position=position_dodge(0.2))+
  geom_point(aes(Temp,propinf_all, colour=Milkweed, shape=Milkweed),
               position=position_dodge(0.2), cex=3)+
  # geom_dotplot(aes(colour=Milkweed, fill=Milkweed), binwidth = 0.03, binaxis='y', stackdir='center',dotsize = 0.5, position=position_dodge(0.2))+
  labs(fill = "Milkweed", x = "temperature treatment", y = "infection probability")+
  scale_fill_manual(values = c("darkgreen","darkorange"))+
  scale_colour_manual(values = c("darkgreen","darkorange"))


```

## Spore load
```{r}
##log10 transformed
m0<-lm(LogOE ~ 1, data = mw.inoc1)
m1<-lm(LogOE ~ Temp + Milkweed + OE_strain + Sex + PupalMass, data = mw.inoc1)
m2<-lm(LogOE ~ Temp*Milkweed + Sex + PupalMass + OE_strain, data = mw.inoc1)

AICc(m0,m1,m2)#model with interaction is best

summary(m2) #interaction signif: p= 0.0393; OE strain signif p=0.03...

em <- emmeans(m2, specs ="Temp", by ="Milkweed")
contrast(em, "pairwise", adjust = "Tukey")

emmip(m2, Milkweed ~ Temp, CIs = TRUE, type = "response") +
  theme_bw()+
  scale_colour_manual(values = c("darkgreen","orange"))


plot(m2) #not very normal qq plot; might need to consider another transformation or dist

##########################################################################
#cube root transformed
m0<-lm((OE_5ml)^(1/3) ~ 1, data = mw.inoc1)
m1<-lm((OE_5ml)^(1/3) ~ Temp + Milkweed + OE_strain + Sex + PupalMass, data = mw.inoc1)
m2<-lm((OE_5ml)^(1/3) ~ Temp*Milkweed + OE_strain + Sex + PupalMass, data = mw.inoc1)


AICc(m0,m1,m2)
#neither is better than null

plot(m2)# better qq plot
hist(m2$residuals)

summary(m2)
tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

# em <- emmeans(m2, specs ="Temp", by ="Milkweed")
# contrast(em, "pairwise", adjust = "Tukey")
# 
# emmip(m2, Milkweed ~ Temp, CIs = TRUE, type = "response") +
#   theme_bw()+
#   scale_colour_manual(values = c("darkgreen","orange"))


##########################################

ggplot(mw.inoc1, aes(x=Temp, y=LogOE, fill=Milkweed)) + 
  geom_boxplot(alpha=0.6)+
  geom_jitter(position = position_jitterdodge(jitter.height = 0.01, 
              jitter.width = 0.3), alpha=0.7, aes(colour=Milkweed))+
  labs(fill = "Milkweed", x = "temperature treatment", y = "parasite load (log10(spores/monarch))")+
  theme_bw()+
  scale_fill_manual(values = c("darkgreen", "orange"))+
  scale_colour_manual(values = c("darkgreen","orange"))


ggplot(mw.inoc1, aes(x=Temp, y=(OE_5ml)^(1/3), fill=Milkweed)) +
  geom_boxplot(alpha=0.6, outliers = FALSE)+
  geom_jitter(position = position_jitterdodge(jitter.height = 0.01,
              jitter.width = 0.3), alpha=0.7, aes(colour=Milkweed))+
  labs(fill = "Milkweed", x = "temperature treatment", y = "parasite load (cube root(spores/monarch))")+
  theme_bw()+
  scale_fill_manual(values = c("darkgreen", "orange"))+
  scale_colour_manual(values = c("darkgreen","orange"))


# Clean data for plotting
mw.inoc1_clean <- mw.inoc1 %>%
  filter(!is.na(OE_5ml))  # Remove NA values for plotting

y_breaks <- c(0, 1000, 50000, 500000, 1000000, 2500000)

# adjust; cube root of 0 is 0
breaks_transformed <- c(0, y_breaks[-1]^(1/3))
labels_formatted <- scales::comma(y_breaks)

ggplot(mw.inoc1_clean, aes(x=Temp, 
                           y=ifelse(OE_5ml==0,0,(OE_5ml)^(1/3)), 
                           fill=Milkweed)) + 
  geom_boxplot(alpha=0.6, outliers = FALSE)+
  geom_jitter(position = position_jitterdodge(jitter.height = 0.01, 
              jitter.width = 0.3), alpha=0.7, 
              aes(colour=Milkweed, shape=Milkweed))+
  labs(fill = "Milkweed", x = "temperature treatment", y = "parasite load (spores/monarch)")+
  theme_bw()+
  # scale_y_continuous(
  #   breaks = y_breaks^(1/3),
  #   labels = scales::comma(y_breaks)) +
  scale_y_continuous(
    breaks = breaks_transformed,
    labels = labels_formatted) +
  theme(axis.text.y = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("darkgreen", "orange"))+
  scale_colour_manual(values = c("darkgreen","orange"))


################################
#performance::check_heteroscedasticity(m2)
#shapiro.test(m2$residuals)

```

##Adult longevity - spore load relationship
```{r}
mw2021.1<-filter(mw2021, !is.na(PupalMass))
mw2021.1<-filter(mw2021.1, !is.na(Sex))
mw2021.1$LogOE<-ifelse(!is.na(mw2021.1$LogOE),mw2021.1$LogOE,0)
mw2021.1$OE_5ml<-ifelse(!is.na(mw2021.1$OE_5ml),mw2021.1$OE_5ml,0)

##########################
#log10 of longevity with guassian distribution
m0<-lm(log10(AdultLongevity) ~ 1, data = mw2021.1)
m1<-lm(log10(AdultLongevity) ~ sqrt(OE_5ml) + InfectionStatus + Sex + PupalMass + Temp + Milkweed, data = mw2021.1)
m2<-lm(log10(AdultLongevity) ~ sqrt(OE_5ml)*Temp + Milkweed + InfectionStatus + Sex + PupalMass, data = mw2021.1)
m3<-lm(log10(AdultLongevity) ~ sqrt(OE_5ml)*Milkweed + Temp + InfectionStatus + Sex + PupalMass, data = mw2021.1)


AICc(m0,m1,m2,m3)#model with OE-temp interaction signif better than null with just main effects

summary(m2) #OE load, pupal mass, milkweed, and temp*OEload interaction signif
tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

#model check
plot(m2)
shapiro.test(resid(m2)) #normal residuals
bptest(m2) #homogeneity
overdisp_fun(m2) #not overdispersed

ggplot(mw2021.1, aes(x=sqrt(OE_5ml), y=log10(AdultLongevity))) +
  geom_smooth(aes(colour=Temp, fill=Temp), method = "lm")+
  geom_point(aes(colour=Temp))+
  labs(fill = "Temp", x = "parasite load (sqrt(spores/monarch))", y = "adult starvation resistance (log10(days))")+
  theme_bw()+
  scale_colour_manual(values = c("#26185F", "#18BDB0"))+
  scale_fill_manual(values = c("#26185F", "#18BDB0"))+
  scale_shape_manual(values = c(25,22))

#adjust axes:
x_breaks <- c(0, 50000, 500000, 1000000, 2500000) #
x_breaks_transformed <- c(0, sqrt(x_breaks[-1]))

y_breaks <- c(0,5,10,15,20,25,30)

ggplot(mw2021.1, aes(x=sqrt(OE_5ml), y=(AdultLongevity))) + 
  geom_smooth(aes(colour=Temp, fill=Temp), method = "lm")+
  geom_point(aes(colour=Temp, fill = Temp, shape=Temp))+
  labs(fill = "Temp", x = "parasite load (spores/monarch)", y = "adult starvation resistance (days)")+
  theme_bw()+
  scale_x_continuous(
    breaks = x_breaks_transformed,
    labels = scales::comma(x_breaks)
  ) +
  scale_y_log10(
    breaks = y_breaks,
    labels = y_breaks
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_colour_manual(values = c("#26185F", "#18BDB0"))+
  scale_fill_manual(values = c("#26185F", "#18BDB0"))+
  scale_shape_manual(values = c(25,22))


```

##########################################################################

**Plant responses to temperature**

## mean C:N ratio in ambient vs elevated
```{r}
m0<-lm(CtoN ~ 1, data=mw2021)
m1<-lm(CtoN ~ Temp + Milkweed, data=mw2021)
m2<-lm(CtoN ~ Temp*Milkweed, data=mw2021)

AICc(m0,m1,m2) 

#interaction is marginally significant
summary(m2) #temp, milkweed, and the interaction all predict C:N
tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

plot(m2)

#summary table for C:N ratio
CtoN_treatments <-
  mw2021 %>% 
  group_by(Temp, Milkweed) %>% 
  dplyr::summarise(mean_CtoN = mean(CtoN, na.rm=T),
                   se_CtoN= sd(CtoN, na.rm = T)/sqrt(n()))

ggplot(CtoN_treatments, aes(x=Temp, y=mean_CtoN, fill=Milkweed)) + 
  geom_bar(stat="identity", 
           position=position_dodge(), alpha=0.8) +
  geom_errorbar(aes(ymin=mean_CtoN-se_CtoN, ymax=mean_CtoN+se_CtoN), width=.2,
                 position=position_dodge(.9))+
  labs(x = "temperature treatment", y = "C:N")+
  theme_bw()+
  scale_fill_manual(values = c("darkgreen","orange"))

```

## mean cardenolide concentration and polarity in tropical plants (ambient vs. elevated)
```{r}
card_am_el_tropical<-
  cardenolides  %>% 
  group_by(Temp) %>% 
  dplyr::summarise( mean_concentration= mean(total.card.conc, na.rm = T),
                    se = sd(total.card.conc, na.rm = T)/sqrt(n()),
                    mean_polarity = mean(sum.of.polarity, na.rm=T),
                    se_pol = sd(sum.of.polarity, na.rm = T)/sqrt(n()))

################
#Go forward excluding 0s
cardenolides_nozero_tropical<-filter(cardenolides, total.card.conc>0)

card_am_el_tropical<-
  cardenolides_nozero_tropical  %>% 
  group_by(Temp) %>% 
  dplyr::summarise( mean_concentration= mean(total.card.conc, na.rm = T),
                    se = sd(total.card.conc, na.rm = T)/sqrt(n()),
                    N_sampes = n(),
                    mean_polarity = mean(sum.of.polarity, na.rm=T),
                    se_pol = sd(sum.of.polarity, na.rm = T)/sqrt(n()))


ggplot(card_am_el_tropical, aes(x=factor(Temp), y=mean_concentration, fill=Temp)) + 
  geom_bar(stat="identity", 
           position=position_dodge(), alpha=0.8) +
  geom_errorbar(aes(ymin=mean_concentration-se, ymax=mean_concentration+se), width=.2,
                 position=position_dodge(.9))+
  theme_bw()+
  labs(x = "temperature treatment", y = "cardenolide concentration (mg/g dry mass)")+
  scale_fill_manual(values = c("#26185F","#18BDB0"))


ggplot(card_am_el_tropical, aes(x=factor(Temp), y=mean_polarity, fill=Temp)) + 
  geom_bar(stat="identity", 
           position=position_dodge(), alpha=0.8) +
  geom_errorbar(aes(ymin=mean_polarity-se_pol, ymax=mean_polarity+se_pol), width=.2,
                 position=position_dodge(.9))+
  labs(x = "Temperature (C)", y = "Average polarity ()")+
  scale_fill_manual(values = c("#26185F","#18BDB0"))

#################################################

norm<-cardenolides %>%
  group_by(Temp) %>%
  rstatix::shapiro_test(total.card.conc) #not normally dist - use wilcox

var<-cardenolides %>% rstatix::levene_test(total.card.conc ~ Temp) #equal variance

#ttest<-t.test(cardenolides$total.card.conc ~ cardenolides$Temp)
#difference in mean cardenolide concentration is not significantly different from 0
wtest<-wilcox.test(cardenolides$total.card.conc ~ cardenolides$Temp, alternative = "two.sided") #difference in card concentration not different from 0
wtest

norm<-cardenolides %>%
  group_by(Temp) %>%
  rstatix::shapiro_test(sum.of.polarity) #not normally dist - use wilcox

var<-cardenolides %>% rstatix::levene_test(sum.of.polarity ~ Temp) #equal variance

#ttest<-t.test(cardenolides$sum.of.polarity ~ cardenolides$Temp)
#difference in mean polarity is not significantly different from 0
wtest<-wilcox.test(cardenolides$sum.of.polarity ~ cardenolides$Temp, alternative = "two.sided")
wtest
#difference in mean polarity is not significantly different from 0
```